Four reasons your build is slow

### 2019-01-17 paper.tex
committing to pull

\section{Analysis tool}
\label{sec:analysis}
In this section we describe the tools we use to analyse make builds, diagnose problems related
to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
discuss the modified version of GNU Make \cite{} we use to record enough information about a
project's build to reconstruct the build graph during a post-mortem analysis.
In section \ref{sec:straceinstr} we discuss how we use a slightly modified version of strace to
record all of the system calls made during the project's build and then use the system calls
to determine the real dependencies of the build. In section \ref{sec:revisegraphs} we discuss
how to combine the information from these two tools to transform a build graph into
a new graph with potentially more parallelism.

@@

To collect the desired level of information from a run of GNU Make \cite{}, we run it with the
debug level of 'verbose', use our own wrappers, which print additional information, and use
a modified version of make which prints directory and timing information for each shell command run.
All of this information is written to a file and later used to reconstruct
the build graph of a successful build.  The reconstructed graph contains all of the shell commands
executed during the build as well as the time each one took to complete.  With the structure of the
build and the timing information the work and span of the build can be calculated. Work is the total
time to execute every shell command in the build graph, and span is the longest time to execute any
sequence of shell commands which must execute sequentially.  For example, the span of a single
make rule is, the longest time to execute a dependency plus the sum of the times to execute each
recipe.  Parallelism is defined as \begin{math} \frac{span}{work} \end{math}, therefore to increase
the parallelism of a project the span must be decreased; making span an important place to focus
performance improvements.

@@

To extend the analysis performable on build graphs, strace \cite{} is used in
conjunction with our enhanced version of GNU Make.  Strace can capture and record the system
calls of every process launched during a make build, allowing us to later map directly between
a shell command launched by make and the system calls it ran.  Strace can write the system call
information for each pid used to its own file.  This is a good way to map directly from a
shell command to a set of system calls, except that Linux may reuse pids during the course of
a build, which causes strace, by default, overwrite these files.  To solve this problem and
maintain a 1 to 1 mapping between processes launched by make and system call information,
we modified strace to append a timestamp to each file extension along with the pid.

@@

Facebook's zstd project \cite{} has a build that exhibits this property.  When we used our tool to analyse
the span of zstd's build we saw that it is dominated by a single call to cc; see figure \ref{code:cc1}.

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  cc \$(FLAGS)
  ../lib/common/debug.c
  ../lib/common/entropy\_common.c
  \textrm{\emph{25 more .c files}}
  zstdcli.o fileio.o bench.o
  datagen.o dibio.o -o zstd
\end{Verbatim}
\caption{A long running cc command which was responsible for ??\% of zstd's span.  It compiled and linked 25 .c files.}
\label{code:cc1}
\end{figure}

If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
see figure \ref{code:cc2}.
\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  cc \$(FLAGS)
  common/debug.c
  common/entropy\_common.c
  \textrm{\emph{28 more .c files}}
  -shared -fPIC -fvisibility=hidden
  -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
\end{Verbatim}
\caption{A long running cc command which was responsible for ??\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
This command compiled and linked 30 files.}
\label{code:cc2}
\end{figure}

If neither of these calls are members of the build's span, then our tool predicts the span would be
approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
This is about an 80\% reduction in the build's span, and the parallelism improves from 2.95 to ??.

To test this we manually replaced the cc command in figure \ref{code:cc1} with the code in figure
\ref{code:cc3}.  The original command is replace by a new target for each object file and target
that can build each object file in parallel and link them together afterwards.

@@

Similarly, the code in figure \ref{code:cc2} is replaced by the code in figure \ref{code:cc4}.  Once
again each object file is explicitly declared as its own target and a target is declared that can
build these object files in parallel and then link them together afterwards.

@@

Makefiles are a collection of rules which describe how to build some targets.  Each rule consists
of a a series of prerequisites and a series of recipes which are run to build the target.  The
prerequisites can be built in parallel but the recipes are run in the order they are listed.  So,
the author of a makefile can affect the potential parallelism of the build by how they write their
makefile.

A makefile which is too sequential will have a longer than necessary span.  In the previous
section we talked about projects having a long span because it was dominated by a single
long running shell command.  Another possiblility is the span being made up of many shell
commands; none of which dominate.  In the previous section we talked about zstd, and if we
rewrite the makefiles so the long running commands are made up of many sequential compilation
and then linking commands we end up with a build that has the same work and span as the original.

@@

\begin{figure}[t]
  \includegraphics[width=0.5\textwidth]{llvm-speedup}
  \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
    to the potential speedups our tool predicted using the work and span calculations}
  \label{fig:llvm}
\end{figure}

Some projects may exhibit ample potential parallelism but may not achieve that potential
parallelism in practice.  LLVM \cite{} is the largest
project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
These calculations where made from a build of LLVM that was run without strace.

@@

Google's protocol buffers project \cite{} exhibited this behavior.  It had a relatively
long predicted span of ?? seconds, which was dominated by a call to g++ which compiles
a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
protocol buffers' build, we learn that if the code in fig. \ref{code:g++1} wasn't part of the span,
the new span would be ?? and would be dominated by the code in fig. \ref{code:g++2}.  And when that
code isn't part of the span, we learn that the new span is again dominated by another
g++ call in the code in fig. \ref{g++3}.  If none of these g++ compilations are in the span,
the new span of the project is roughly half the original span.

@@

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/descriptor.lo
  google/protobuf/descriptor.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
40 seconds to run.}
\label{code:g++1}
\end{figure}

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/descriptor.pb.lo
  google/protobuf/descriptor.pb.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
27 seconds to run.}
\label{code:g++2}
\end{figure}

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/compiler/cpp/cpp\_message.lo
  google/protobuf/compiler/cpp/cpp\_message.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
22.5 seconds to run.}
\label{code:g++3}
\end{figure}

@@

One of the limitations of this work is that currenlty to collect the necessary data to reconstruct
the build graph, the project must be built using a single thread.  This is a burden for projects
that are very long running, such as LLVM.  Additionally, building with strace causes an additional
slowdown and inflates the running time of each shell command, which makes the raw numbers
reported by a build graph created with strace on, inaccurate.  In the future for performance sake
it would be good to collect the necessary information while allowing make to run in parallel with
any number of processes.  It would also be good to know how much strace perturbs the recorded
time of any given shell command so both the tool with and without strace don't need to be run.

@@

>>>>

\section{Analysis tool}
\label{sec:analysis}
In this section we describe the tools we use to analyse make builds, diagnose problems related
to parallelism, and identify places for improvement. In section \ref{sec:makeanalysis} we
discuss the modified version of GNU Make \cite{gnumake4.2} we use to record enough information about a
project's build to reconstruct the build graph during a post-mortem analysis.
In section \ref{sec:straceinstr} we discuss how we use a slightly modified version of strace to
record all of the system calls made during the project's build and then use the system calls
to determine the real dependencies of the build. In section \ref{sec:revisegraphs} we discuss
how to combine the information from these two tools to transform a build graph into
a new graph with potentially more parallelism.

@@

To collect the desired level of information from a run of GNU Make \cite{gnumake4.2}, we run it with the
debug level of 'verbose', use our own wrappers, which print additional information, and use
a modified version of make which prints directory and timing information for each shell command run.
All of this information is written to a file and later used to reconstruct
the build graph of a successful build.  The reconstructed graph contains all of the shell commands
executed during the build as well as the time each one took to complete.  With the structure of the
build and the timing information the work and span of the build can be calculated. Work is the total
time to execute every shell command in the build graph, and span is the longest time to execute any
sequence of shell commands which must execute sequentially.  For example, the span of a single
make rule is, the longest time to execute a dependency plus the sum of the times to execute each
recipe.  Parallelism is defined as \begin{math} \frac{span}{work} \end{math}, therefore to increase
the parallelism of a project the span must be decreased; making span an important place to focus
performance improvements.

@@

To extend the analysis performable on build graphs, strace \cite{strace4.25} is used in
conjunction with our enhanced version of GNU Make.  Strace can capture and record the system
calls of every process launched during a make build, allowing us to later map directly between
a shell command launched by make and the system calls it ran.  Strace can write the system call
information for each pid used to its own file.  This is a good way to map directly from a
shell command to a set of system calls, except that Linux may reuse pids during the course of
a build, which causes strace, by default, overwrite these files.  To solve this problem and
maintain a 1 to 1 mapping between processes launched by make and system call information,
we modified strace to append a timestamp to each file extension along with the pid.

@@

Facebook's zstd project \cite{zstd1.3.7} has a build that exhibits this property.  When we used our tool to analyse
the span of zstd's build we saw that it is dominated by a single call to cc; see fig. \ref{code:cc1}.

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  cc \$(FLAGS)
  ../lib/common/debug.c
  ../lib/common/entropy\_common.c
  \textrm{\emph{25 more .c files}}
  zstdcli.o fileio.o bench.o
  datagen.o dibio.o -o zstd
\end{Verbatim}
\caption{A long running cc command which was responsible for about 97\% of zstd's span.  It compiled and linked 27 .c files.}
\label{code:cc1}
\end{figure}

If this call is no longer part of the span, the new span is similarly dominated by another single call to cc;
see fig. \ref{code:cc2}.
\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  cc \$(FLAGS)
  common/debug.c
  common/entropy\_common.c
  \textrm{\emph{28 more .c files}}
  -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
\end{Verbatim}
\caption{A long running cc command which was responsible for aproximately 100\% of zstd's span when \ref{code:cc1} was not a membe rof the span.
This command compiled and linked 30 files.}
\label{code:cc2}
\end{figure}

If neither of these calls are members of the build's span, then our tool predicts the span would be
approximately 7.2 seconds rather than 39 seconds, as with the original span, or 37 seconds, as with the second span.
This is about an 80\% reduction in the build's span, and the parallelism improves from 2.98 to 16.

To test this we manually replaced the cc command in fig. \ref{code:cc1} with the code in fig.
\ref{code:cc3}.  The original command is replace by a new target for each object file and target
that can build each object file in parallel and link them together afterwards.

@@

Similarly, the code in fig. \ref{code:cc2} is replaced by the code in fig. \ref{code:cc4}.  Once
again each object file is explicitly declared as its own target and a target is declared that can
build these object files in parallel and then link them together afterwards.

@@

Makefiles are a collection of rules which describe how to build targets \cite{gnumakemanual}.
Each rule consists
of a a series of prerequisites and a series of recipes which are run to build the target.  The
prerequisites can be built in parallel, but the recipes must be run in the order they are listed.  So,
the author of a makefile can directly impact the potential parallelism of the build by
how they organize their makefile.

In the previous section we discussed zstd's build, which turned out to be unnecessarily sequential
because two single shell calls which compiled and linked many files at once could be broken up
into many parallel compilation jobs which could then be linked after.  These two commands could
have been written as follows in fig. \ref{code:cc5} and fig. \ref{code:cc6} respectively, instead of as they were
originally in fig. \ref{code:cc1} and fig. \ref{code:cc2} and would have resulted in the same
build performance.

\begin{figure}[H]
  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
cc \$(FLAGS) ../lib/common/debug.c
   -o ../lib/common/debug.o
cc \$(FLAGS) ../lib/common/entropy\_common.c
   -o ../lib/common/entropy\_common.o
\textrm{\emph{25 more recipes; one for each .c file that needs to be compiled}}
cc \$(FLAGS) \$(OBJ\_FILES) -o zstd
\end{Verbatim}
  \caption{An alternative to the code in fig. \ref{code:cc1}. Each file is compiled individually
    as its own recipe, followed by a recipe to link the object files together.  This has the same
    performance as the original command.}
\label{code:cc5}
\end{figure}

\begin{figure}[H]
  \begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
cc \$(FLAGS) common/debug.c -o common/debug.o1
cc \$(FLAGS) common/entropy\_common.c
   -o common/\_common.o1
\textrm{\emph{28 more recipes; one for each .c file that needs to be compiled}}
cc \$(FLAGS) \$(OBJ\_FILES) -shared
   -Wl,-soname=libzstd.so.1 -o libzstd.so.1.3.7
\end{Verbatim}
  \caption{An alternative to the code in fig. \ref{code:cc2}. Each file is compiled individually
    as its own recipe, followed by a recipe to link the object files together.  This has the same
    performance as the original command.}
\label{code:cc6}
\end{figure}

@@

\begin{figure}[t]
  \includegraphics[width=0.5\textwidth]{llvm-speedup-w-64}
  \caption{A comparison of the observed speedup when LLVM was built with \emph{make -j <n> for <n> from 1 to 32},
    to the potential speedups our tool predicted using the work and span calculations}
  \label{fig:llvm}
\end{figure}

Some projects may exhibit ample potential parallelism but may not achieve that potential
parallelism in practice.  LLVM \cite{llvm6.0.0} is the largest
project we analysed, with a calculated work of approximately 12584 seconds, or 3.5 hours.
With a predicted span of 305 seconds, or 5 minutes, LLVM has a parallelism of about 41.
These calculations where made from a build of LLVM that was run without strace.

@@

Google's protocol buffers project \cite{protobufs3.6.1} exhibited this behavior.  It's work was calculated
to be about 16.7 minutes and it had a relatively
long predicted span of 43.7 seconds, which was dominated by a call to g++ compiling
a single file.  When the analysis applied in section \ref{sec:expensive} is applied to
protocol buffers' build, we learn that if the code in fig. \ref{code:g++1} wasn't part of the span,
the new span would be about 31s and would be dominated by the code in fig. \ref{code:g++2}.
And when that
code isn't part of the span, we learn that the new span is again dominated by another
g++ call in the code in fig. \ref{g++3}.  If none of these g++ compilations are in the span,
the new span of the project is about 20s, roughtly half of the original span.

@@

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/descriptor.lo
  google/protobuf/descriptor.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
40 seconds to run and was 90\% of the span.}
\label{code:g++1}
\end{figure}

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/descriptor.pb.lo
  google/protobuf/descriptor.pb.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
27 seconds to run and was about 87\% of the span.}
\label{code:g++2}
\end{figure}

\begin{figure}[H]
\begin{Verbatim}[commandchars=\\\{\},codes={\catcode`$=3\catcode`^=7\catcode`_=8},fontsize=\small,numbers=left,xleftmargin=5mm]
  g++ \$(FLAGS)
  -o google/protobuf/compiler/cpp/cpp\_message.lo
  google/protobuf/compiler/cpp/cpp\_message.cc
\end{Verbatim}
\caption{A simplification of the real command run by make; compiles a single file and took about
22.5 seconds to run and was about 90\% of the span.}
\label{code:g++3}
\end{figure}

@@

One of the limitations of this work is that currently to collect the necessary data to reconstruct
the build graph, the project must be built using a single thread.  This is a burden for projects
that are very long running, such as LLVM.  Additionally, building with strace causes an additional
slowdown and inflates the running time of each shell command, which makes the raw numbers
reported by a build graph created with strace on, inaccurate.  In the future for performance sake
it would be good to collect the necessary information while allowing make to run in parallel with
any number of processes.  It would also be good to know how much strace perturbs the recorded
time of any given shell command so both the tool with and without strace don't need to be run.

@@

\bibliography{citations}
\bibliographystyle{plain}
